---
title: "Sample Kernel for HR Dataset v13"
author: "Dr. Rich Huebner"
date: "10/4/2019"
output: 
  html_document: 
    highlight: tango
    theme: cerulean
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(magrittr)
library(bnlearn)
library(rpart)
library(rpart.plot)
library(rpart.utils)
library(factoextra)
```


## Sample Kernel for HR Data Set v13



## Structure of Data


```{r}
# Read data and show structure

df <- read.csv("HRDataset_v13.csv", sep=",", header=TRUE)
str(df)

```

## A bit of feature engineering and cleanup

```{r}
df$Position2  <- as.character(df$Position)
sl <- sort(unique(df$Position2))
for (s in 1:length(sl)) df$Position2[df$Position == sl[s]] <- LETTERS[s]    # A, B, C, D.
df$Position2 <- as.factor(df$Position2)

#simlar thing for department
df$Department2 <- as.character(df$Department)
dl <- sort(unique(df$Department2))
for (s in 1:length(dl)) df$Department2[df$Department == dl[s]] <- LETTERS[s]   # A, B, C...
df$Department2 <- as.factor(df$Department2)



# Cleanup so that the visualizations are a little clearer
df$RaceCd2[df$RaceDesc == 'American Indian or Alaska Native'] <- "AmerInd"
df$RaceCd2[df$RaceDesc == 'White'] <- "White"
df$RaceCd2[df$RaceDesc == 'Black or African American'] <- "Black"
df$RaceCd2[df$RaceDesc == 'Asian'] <- "Asian"
df$RaceCd2[df$RaceDesc == 'Hispanic'] <- "Hisp"
df$RaceCd2[df$RaceDesc == 'Two or more races'] <- ">2"

```



## Problems with Attendance?


```{r}

ggplot(df) +
 aes(x = RaceCd2, y = DaysLateLast30, fill = Department) +
 geom_boxplot() +
 scale_fill_hue() +
 labs(title = "Who's Late?", subtitle = "An analysis of who is late") +
 theme_minimal()

```

## Performance Score Analysis By Department

```{r}

ggplot(df) +
 aes(x = Department2, y = EngagementSurvey) +
 geom_boxplot(fill = "#33a02c") +
 labs(title = "Engagement Survey Scores ", subtitle = "By Department, By Gender") +
 theme_pander() +
 facet_wrap(vars(Sex))

```

## Pay Across Departments

```{r}

df <- df %>%
 filter(!(TermReason %in% ""))

ggplot(df) +
 aes(x = Department, y = PayRate) +
 geom_boxplot(fill = "#ffccee") +
 labs(title = "Pay Across Departments") +
 theme_economist_white()

```

Not surprising that Software Engineering and Sales people are paid pretty high. The people that work on the production floor, however, are the 
lowest paid within the business.


## Pay Distribution

```{r}
ggplot(df) +
 aes(x = PayRate) +
 geom_histogram(bins = 30L, fill = "#6633ff") +
 labs(title = "Pay Rate Distribution - Histogram") +
 theme_pander()
```


## Engagement Survey Results 

```{r}
# Engagement Survey vs. Payrate -- anything there?
# Figure out average engagement survey score per recruitment source

k <- df %>%
  dplyr::group_by(RecruitmentSource) %>%
  summarize(mean_engagement = mean(EngagementSurvey, na.rm = TRUE))

ggplot(k, aes(x=mean_engagement, y=RecruitmentSource, label = round(mean_engagement, digits = 2))) +
 geom_text(size = 4) +
 labs(title = "Engagement Survey Results vs. Recruitment Source")


p <- df %>%
  dplyr::group_by(Department) %>%
  summarize(mean_engagement = mean(EngagementSurvey, na.rm = TRUE))

ggplot(p, aes(x=mean_engagement, y=Department, label = round(mean_engagement, digits = 2))) +
 geom_text(size = 4) +
 labs(title = "Engagement Survey Results vs. Department")

q <- df%>%
  dplyr::group_by(RaceCd2) %>%
  summarize(mean_engagement = mean(EngagementSurvey, na.rm = TRUE))
ggplot(q, aes(x=mean_engagement, y=RaceCd2, label = round(mean_engagement, digits = 2))) +
 geom_text(size = 4) +
 labs(title = "Engagement Survey Results by Race Code")


```


## Number of Staff

```{r}
 s <- df %>%
  dplyr::group_by(Position) %>%
  summarize(num = n())

 s$Position <- as.character(s$Position)
 
 
 ggplot(s, aes(x=reorder(Position, num), y=num)) +
   geom_text(aes(y=num, label=num)) +
   theme_bw()+
   theme(axis.text.x = element_text(angle=60, hjust=1),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.x = element_line(color="Blue", linetype="dotted")) +
   labs(title = "Number of Employees per Position")
 
 
 # num of employees per department
 t <- df %>%
   dplyr::group_by(Department) %>%
   summarize(num = n())
 
  ggplot(t, aes(x=reorder(Department, num), y=num)) +
   geom_text(aes(y=num, label=num)) +
   theme_bw()+
   theme(axis.text.x = element_text(angle=45, hjust=1),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.x = element_line(color="Blue", linetype="dotted")) +
   labs(title = "Number of Employees per Department")

```


## Relationships Between Pay and Engagement Survey

```{r}

ggplot(df) +
 aes(x = PayRate, y = EngagementSurvey) +
 geom_point(size = 1.56, colour = "#08519c") +
 #geom_smooth(span = 0.75) +
 theme_light()

```

## Clustering Approach - Unsupervised Learning Method

```{r}

s2<- df %>%
        select(c(PayRate, EngagementSurvey, EmpSatisfaction, SpecialProjectsCount))
s2_scaled <- scale(s2, center = TRUE, scale = TRUE)
s2_scaled <- na.omit(s2_scaled)

k_model <-  kmeans(s2_scaled, iter.max = 20, nstart = 20, centers = 2)
p1 <- fviz_cluster(k_model, data = s2_scaled, geom = "point", repel = FALSE, ellipse = TRUE,
                   palette = "Set1", ellipse.type = "norm", show.clust.cent = TRUE,
                   ggtheme = theme_minimal(),
                   main = "kMeans Clustering Plot")
p1

```

kMeans clustering only explains 61% of the variability of these four variables.
The structure of this data set could be explored further with more advacned clustering and unsupervised techniques.

## Decision Tree Example

```{r}
ctl_tree <- rpart.control(minsplit = 20, maxdepth = 3)
tree1 <- rpart( Termd ~ PayRate + Sex + MaritalDesc + RaceDesc + Department + PerformanceScore + SpecialProjectsCount, 
                  data=df, method = "class", control = ctl_tree)      #, na.action=na.pass)
rpart.plot(tree1, type=5, box.palette="RdBu", shadow.col="gray", nn=TRUE, main=(paste("Decision Tree Example")))

```

Next, we display the Decision Tree Rules:

```{r}
  rattle::asRules(tree1)
```